---
title: "Hepatectomy: Right vs Left Robotic"
author: "Oscar J. Ponce, Eddy Lincango & Luis A. Figueroa"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(dirname(inputFile),'/README.md')) })
output: github_document
---
<i> Analysis and visualization developed by [Oscar J. Ponce](https://twitter.com/PonceOJ), [Eddy Lincango](https://twitter.com/DoctorEddyMD) and [Luis A. Figueroa](https://twitter.com/LuisFig1706) </i>

The purpose of this site is to make our meta-analysis as transparent as possible. All information or documents used, from the protocol to the final publication, will be available here. If further information is needed, please do not hesitate to contact us.

```{r setup, include=FALSE}
library(metafor)
library(tidyverse)
library(forestplot)
library(rmarkdown)

knitr::opts_chunk$set(fig.path = "Cont Output Figures/", dev='svg')
```



```{r functions_md_or, echo=FALSE, message=FALSE, warning=FALSE}
md <- function(database){
  
  db <- database
  db <- escalc(measure="MD", 
               m1i=mean_post1, m2i = mean_post2, 
               sd1i  = sd_post1, sd2i = sd_post2, 
               n1i = n1, n2i= n2, data=database)
  db$vi <- ifelse(is.na(db$vi), 
               ((db$mdul_post-db$mdll_post)/((2*abs(qt(0.05/2, db$total-1)))^2)), db$vi)
  db <- db[order(db$yi),]
  db <- summary(db)
  
  db$md <- paste(formatC(db$yi, format='f', digits =1)," ",
                 "(",formatC(db$ci.lb, format='f', digits =1),
                 ",",formatC(db$ci.ub, format='f', digits=1),")") 
  db$postmean1 <- paste(formatC(db$mean_post1, format='f', digits=1),'(', formatC(db$sd_post1, format='f', digits=1),')')
  db$postmean2 <- paste(formatC(db$mean_post2, format='f', digits=1),'(', formatC(db$sd_post2, format='f', digits=1),')')
  
  ma <- rma(yi, vi, measure='MD', data=db, method='REML')
  
  db$w <- paste(formatC(weights(ma), format='f', digits = 1),'%')
  db$weights <- weights(ma)
  
  list(pre = db, ma = ma)
}


table_md <- function(analysis, nstudies, int, comp, outcome, col, 
                     follow=FALSE){
   ifelse(nstudies>1,
        (b <- cbind( 
          c("Author", analysis$pre$author, 
            paste("Overall Mean Difference for", analysis$ma$k, "studies"),
            paste("(Tau^2 = ", (formatC(analysis$ma$tau2, digits=2, format="f")), ", df = ", 
                  (analysis$ma$k - analysis$ma$p),
                  ", p ", (ifelse(analysis$ma$QEp < 0.001, 
                                  paste("< 0.001"),
                                  paste("= ", formatC(analysis$ma$QEp, digits=3, format="f")))),
                  "; ", "I^2", " = ", (formatC(analysis$ma$I2, digits=1, format="f")), "%)")),
        c(paste(int), analysis$pre$n1, sum(analysis$pre$n1),NA),
        c(paste(outcome, '\n', col), analysis$pre$postmean1, NA, NA),
        c(paste(comp), analysis$pre$n2,sum(analysis$pre$n2), NA),
        c(paste(outcome, '\n', col), analysis$pre$postmean2, NA, NA),
        c("Mean Difference (95% CI)", analysis$pre$md, 
            paste(formatC(analysis$ma$b, format='f', digits =1), 
                  " (",formatC(analysis$ma$ci.lb, format='f', digits=1),
                  ",", formatC(analysis$ma$ci.ub, format='f', digits=1), ")"), NA),
        c("Weight (%)", analysis$pre$w, NA, NA))),
        (b <- cbind( 
        c("Author", analysis$pre$author),
        c(paste(int), analysis$pre$n1),
        c(paste(outcome,'\n', col), analysis$pre$postmean1),
        c(paste(comp), analysis$pre$n2),
        c(paste(outcome, '\n',col), analysis$pre$postmean2),
        c("Mean Difference (95% CI)", analysis$pre$md))))

  ifelse(follow==TRUE, 
          b <- cbind(b, c(paste('Follow-up'), analysis$pre$followup, NA)), NA)
  ifelse(follow==TRUE,
          b <- b[,c(1,ncol(b),2:(ncol(b)-1))], NA)
  
  
  b <- as_tibble(b, .name_repair = "unique") 
  b <- b %>% add_row(.before = 2)
  b <- b %>% add_row(.before = 2)
  b <- b %>% add_row(.after = nrow(b))
  
  ifelse(nstudies>1, b <- b %>% add_row(.before = nrow(b)-2), NA)

  
  ifelse(nstudies > 1,
        (c <- structure(list(
          mean = c(rep(NA, 3), analysis$pre$yi, NA, analysis$ma$b,NA, NA),
          lower = c(rep(NA, 3), analysis$pre$ci.lb, NA, analysis$ma$ci.lb, NA, NA),
          upper = c(rep(NA, 3), analysis$pre$ci.ub, NA, analysis$ma$ci.ub, NA, NA)),
          .Names = c("mean", "lower", "upper"),
          row.names = c(NA, -1L*nrow(b)),
          class = "data.frame")),
        (c <- structure(list(
          mean = c(rep(NA, 3), analysis$pre$yi, NA),
          lower = c(rep(NA, 3), analysis$pre$ci.lb, NA),
          upper = c(rep(NA, 3), analysis$pre$ci.ub, NA)),
          .Names = c("mean", "lower", "upper"),
          row.names = c(NA, -1L*nrow(b)),
          class = "data.frame")))

  c <- as_tibble(c, .name_repair = "unique")
  
  list(b = b, c = c)
} 




plotmd_single_studies <- function(words, numbers,  
                                  xtick, sizes, bolding, aligning, fpPosition) {

  (forestplot(words,
             graph.pos = fpPosition,
             zero = 0,
             numbers,
             new_page = TRUE,
             colgap = unit(5, "mm"),
             hrzl_lines = list('3' = gpar (lwd=1, columns=c(1:(ncol(words)+1)), col="black")),
             lineheight=unit(0.7,'cm'),
             boxsize = sizes,
             line.margin = 2,
             is.summary = bolding,
             align = aligning,
             ci.vertices = TRUE,
             txt_gp = fpTxtGp(label =gpar (cex=0.9), 
                              ticks = gpar(cex = 0.9, fontface="bold"),
                              summary = gpar(cex = 0.9),
                              xlab = gpar(cex=0.9)),
             xticks = xtick,
             xlog=FALSE,
             clip = c(0,1),
             grid = xtick,
             lwd.xaxis = 1,
             lwd.ci = 2.2,
             lwd.zero = 1.5,
             graphwidth = unit(10,"cm"),
             col=fpColors(box="black",line="grey", zero = 'dodgerblue4', axes="grey20", summary="black")))
 
}

```


<h2>Right robotic vs Left robotic Mean Difference (MD) 95%CI</h2>


<details>

<summary><b>Figure A -</b> Blood Loss (mL), Legnth of Hospital Stay (days) and Operative Time (min)</summary>

<br>
```{r overall_graph_REML, echo=FALSE, message=FALSE, warning=FALSE, fig.height = 9.9, fig.width = 15.5}
robotic_cont <- read.csv("data_extraction_continous.csv")
names(robotic_cont)[2] <- "author"
names(robotic_cont)[8] <- "n2"
names(robotic_cont)[9] <- "n1"
names(robotic_cont)[10] <- "mean_post2" #Change for control group
names(robotic_cont)[12] <- "sd_post2"   #Change for control group
names(robotic_cont)[13] <- "mean_post1" #Change for intervention group
names(robotic_cont)[15] <- "sd_post1"   #Change for intervention group


#Blood loss
bl <- subset(robotic_cont, outcome=='blood loss')
bl_md <- md(bl)
tbl_md <- table_md(analysis = bl_md, nstudies = 2, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Total blood loss (bl)', col = 'Mean (SD)', 
                    follow = FALSE)


#Operative time
ot <- subset(robotic_cont, outcome=='Operative time')
ot_md <- md(ot)
tot_md <- table_md(analysis = ot_md, nstudies = 2, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Operative time (min)', col = 'Mean (SD)', 
                    follow = FALSE)

#length of hospital stay
los <- subset(robotic_cont, outcome=='length of hospital stay')
los_md <- md(los)
tlos_md <- table_md(analysis = los_md, nstudies = 8, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Length of Hospital stay (min)', col = 'Mean (SD)', 
                    follow = FALSE)


title <- c('', paste0('Right robotic (N)'), "mean (SD)",
            paste0('Left robotic (N)'), "mean (SD)",
            paste0('Mean Difference (MD) 95%CI'), 
            paste0('Weights'))
  

sub_cont1 <- c('Blood Loss (mL)', rep(NA, 6))
sub_cont2 <- c('Legnth of Hospital Stay (days)', rep(NA, 6))
sub_cont3 <- c('Operative Time (min)', rep(NA, 6))


subtotal2 <- rbind(title, NA,
                   sub_cont1, tbl_md$b[-c(1:3, nrow(tbl_md$b)-3),], NA,
                   sub_cont2, tlos_md$b[-c(1:3, nrow(tlos_md$b)-3),], NA,
                   sub_cont3, tot_md$b[-c(1:3, nrow(tot_md$b)-3),])
subtotal2 <- as.matrix(subtotal2) #Sol: Error in assertAndRetrieveTidyValue(x, lower) : 
#You have not provided an argument and the data frame does not have a 'lower' column: V1, V2, V3, V4, V5, V6, V7

subtotal2_summary <- c(T, F, 
                           T, rep(F, nrow(tbl_md$b[-c(1:3, nrow(tbl_md$b)-3),])-3),T,T,F,F,
                           T, rep(F, nrow(tlos_md$b[-c(1:3, nrow(tlos_md$b)-3),])-3), T, T,F,F,
                           T, rep(F, nrow(tot_md$b[-c(1:3, nrow(tot_md$b)-3),])-3), T, T,F)

sizes_cont <- c(rep(NA,3), 0.006*(bl_md$pre$weights),1,
                rep(NA, 4), 0.03*(los_md$pre$weights),1,
                rep(NA, 4), 0.05*(ot_md$pre$weights),1,
                rep(NA, 2))

subtotal2n <- rbind(NA,NA,
                    NA, tbl_md$c[-c(1:3, nrow(tbl_md$c)-3),], NA,
                   NA, tlos_md$c[-c(1:3, nrow(tlos_md$c)-3),], NA,
                   NA, tot_md$c[-c(1:3, nrow(tot_md$c)-3),])


plot <- forestplot(subtotal2,
             graph.pos = 6,
             zero = 0,
             subtotal2n,
             new_page = TRUE,
             colgap = unit(5, "mm"),
             hrzl_lines = list("2" = gpar (lwd=1, columns=1:8, col="black")),
             lineheight=unit(0.7,'cm'),
             boxsize = sizes_cont,
             line.margin = 2,
             is.summary = subtotal2_summary,
             align = c('l', rep('c', 4), 'l', 'l'),
             ci.vertices = TRUE,
             txt_gp = fpTxtGp(label =gpar (cex=0.9), 
                              ticks = gpar(cex = 0.9, fontface="bold"),
                              summary = gpar(cex = 0.9),
                              xlab = gpar(cex=0.9)),
             xticks = c(-90,-45,0,30,50),
             xlog=FALSE,
             clip = c(-90,-45,30,50),
             grid = c(-90,-45,0,30,50),
             lwd.xaxis = 1,
             lwd.ci = 2.2,
             lwd.zero = 2,
             graphwidth = unit(8,"cm"),
             col=fpColors(box="black",line="grey", zero = 'dodgerblue4', axes="grey20", summary="black"))

plot

```

</details>