---
title: "Hepatectomy: Right vs Left Robotic"
author: "Oscar J. Ponce, Eddy Lincango & Luis A. Figueroa"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(dirname(inputFile),'/README.md')) })
output: github_document
---
<i> Analysis and visualization developed by [Oscar J. Ponce](https://twitter.com/PonceOJ), [Eddy Lincango](https://twitter.com/DoctorEddyMD) and [Luis A. Figueroa](https://twitter.com/LuisFig1706) </i>

The purpose of this site is to make our meta-analysis as transparent as possible. All information or documents used, from the protocol to the final publication, will be available here. If further information is needed, please do not hesitate to contact us.

```{r setup, include=FALSE}
library(metafor)
library(tidyverse)
library(forestplot)
library(rmarkdown)

knitr::opts_chunk$set(fig.path = "Cont Output Figures/", dev='svg')
```



```{r functions_md_or, echo=FALSE, message=FALSE, warning=FALSE}
md <- function(database){
  
  db <- database
  db <- escalc(measure="MD", 
               m1i=mean_post1, m2i = mean_post2, 
               sd1i  = sd_post1, sd2i = sd_post2, 
               n1i = n1, n2i= n2, data=database)
  db$vi <- ifelse(is.na(db$vi), 
               ((db$mdul_post-db$mdll_post)/((2*abs(qt(0.05/2, db$total-1)))^2)), db$vi)
  db <- db[order(db$yi),]
  db <- summary(db)
  
  db$md <- paste(formatC(db$yi, format='f', digits =1)," ",
                 "(",formatC(db$ci.lb, format='f', digits =1),
                 ",",formatC(db$ci.ub, format='f', digits=1),")") 
  db$postmean1 <- paste(formatC(db$mean_post1, format='f', digits=1),'(', formatC(db$sd_post1, format='f', digits=1),')')
  db$postmean2 <- paste(formatC(db$mean_post2, format='f', digits=1),'(', formatC(db$sd_post2, format='f', digits=1),')')
  
  ma <- rma(yi, vi, measure='MD', data=db, method='REML')
  
  db$w <- paste(formatC(weights(ma), format='f', digits = 1),'%')
  db$weights <- weights(ma)
  
  list(pre = db, ma = ma)
}


table_md <- function(analysis, nstudies, int, comp, outcome, col, 
                     follow=FALSE){
   ifelse(nstudies>1,
        (b <- cbind( 
          c("Author", analysis$pre$author, 
            paste("Overall Mean Difference for", analysis$ma$k, "studies"),
            paste("(Tau^2 = ", (formatC(analysis$ma$tau2, digits=2, format="f")), ", df = ", 
                  (analysis$ma$k - analysis$ma$p),
                  ", p ", (ifelse(analysis$ma$QEp < 0.001, 
                                  paste("< 0.001"),
                                  paste("= ", formatC(analysis$ma$QEp, digits=3, format="f")))),
                  "; ", "I^2", " = ", (formatC(analysis$ma$I2, digits=1, format="f")), "%)")),
        c(paste(int), analysis$pre$n1, sum(analysis$pre$n1),NA),
        c(paste(outcome, '\n', col), analysis$pre$postmean1, NA, NA),
        c(paste(comp), analysis$pre$n2,sum(analysis$pre$n2), NA),
        c(paste(outcome, '\n', col), analysis$pre$postmean2, NA, NA),
        c("Mean Difference (95% CI)", analysis$pre$md, 
            paste(formatC(analysis$ma$b, format='f', digits =1), 
                  " (",formatC(analysis$ma$ci.lb, format='f', digits=1),
                  ",", formatC(analysis$ma$ci.ub, format='f', digits=1), ")"), NA),
        c("Weight (%)", analysis$pre$w, NA, NA))),
        (b <- cbind( 
        c("Author", analysis$pre$author),
        c(paste(int), analysis$pre$n1),
        c(paste(outcome,'\n', col), analysis$pre$postmean1),
        c(paste(comp), analysis$pre$n2),
        c(paste(outcome, '\n',col), analysis$pre$postmean2),
        c("Mean Difference (95% CI)", analysis$pre$md))))

  ifelse(follow==TRUE, 
          b <- cbind(b, c(paste('Follow-up'), analysis$pre$followup, NA)), NA)
  ifelse(follow==TRUE,
          b <- b[,c(1,ncol(b),2:(ncol(b)-1))], NA)
  
  
  b <- as_tibble(b) 
  b <- b %>% add_row(.before = 2)
  b <- b %>% add_row(.before = 2)
  b <- b %>% add_row(.after = nrow(b))
  
  ifelse(nstudies>1, b <- b %>% add_row(.before = nrow(b)-2), NA)

  
  ifelse(nstudies > 1,
        (c <- structure(list(
          mean = c(rep(NA, 3), analysis$pre$yi, NA, analysis$ma$b,NA, NA),
          lower = c(rep(NA, 3), analysis$pre$ci.lb, NA, analysis$ma$ci.lb, NA, NA),
          upper = c(rep(NA, 3), analysis$pre$ci.ub, NA, analysis$ma$ci.ub, NA, NA)),
          .Names = c("mean", "lower", "upper"),
          row.names = c(NA, -1L*nrow(b)),
          class = "data.frame")),
        (c <- structure(list(
          mean = c(rep(NA, 3), analysis$pre$yi, NA),
          lower = c(rep(NA, 3), analysis$pre$ci.lb, NA),
          upper = c(rep(NA, 3), analysis$pre$ci.ub, NA)),
          .Names = c("mean", "lower", "upper"),
          row.names = c(NA, -1L*nrow(b)),
          class = "data.frame")))

  c <- as_tibble(c)
  
  list(b = b, c = c)
} 




plotmd_single_studies <- function(words, numbers,  
                                  xtick, sizes, bolding, aligning, fpPosition) {

  (forestplot(words,
             graph.pos = fpPosition,
             zero = 0,
             numbers,
             new_page = TRUE,
             colgap = unit(5, "mm"),
             hrzl_lines = list('3' = gpar (lwd=1, columns=c(1:(ncol(words)+1)), col="black")),
             lineheight=unit(0.7,'cm'),
             boxsize = sizes,
             line.margin = 2,
             is.summary = bolding,
             align = aligning,
             ci.vertices = TRUE,
             txt_gp = fpTxtGp(label =gpar (cex=0.9), 
                              ticks = gpar(cex = 0.9, fontface="bold"),
                              summary = gpar(cex = 0.9),
                              xlab = gpar(cex=0.9)),
             xticks = xtick,
             xlog=FALSE,
             clip = c(0,1),
             grid = xtick,
             lwd.xaxis = 1,
             lwd.ci = 2.2,
             lwd.zero = 1.5,
             graphwidth = unit(10,"cm"),
             col=fpColors(box="black",line="grey", zero = 'dodgerblue4', axes="grey20", summary="black")))
 
}




or <- function(database){
  db <- database
  db <- escalc(measure="OR", ai=e1, ci=e2, n1i=n1, n2i=n2,data=database)
  db <- db[order(db$yi),]
  db <- summary(db)
  
  db$est <-  exp(db$yi)
  db$ci.lb <- exp(db$ci.lb)
  db$ci.ub <- exp(db$ci.ub)
  
  
  db$rate1 <- paste(db$e1,"/",db$n1)
  db$rate2 <- paste(db$e2,"/",db$n2)
  db$or <- paste(formatC(db$est, format='f', digits =2)," ",
                 "(",formatC(db$ci.lb, format='f', digits =2),
                 "-",formatC(db$ci.ub, format='f', digits=2),")")  
  
  ma <- rma(db$yi, db$vi, measure='OR', data=db, method='REML')
  exp <- predict(ma, transf = transf.exp.int)
  
  db$weights <- weights(ma)
  db$w <- paste(formatC(db$weights, format='f', digits = 1),'%')
  
  
  list(pre = db, ma = ma, exp = exp)
}



table_or <- function(analysis, int, comp, outcome, nstudies){
  ifelse(nstudies>1,
  b <- cbind( 
    c("Author", analysis$pre$author, 
      paste("Overall Odds Ratio for", analysis$ma$k, "studies"),
      paste("(Tau^2 = ", (formatC(analysis$ma$tau2, digits=2, format="f")), ", df = ", 
            (analysis$ma$k - analysis$ma$p),
            ", p ", (ifelse(analysis$ma$QEp < 0.001, 
                            paste("< 0.001"),
                            paste("= ", formatC(analysis$ma$QEp, digits=3, format="f")))),
            "; ", "I^2", " = ", (formatC(analysis$ma$I2, digits=1, format="f")), "%)")),
    c(paste(int, '\n', outcome,'/','total n'),
      analysis$pre$rate1, paste(sum(analysis$pre$e1), " / ", sum(analysis$pre$n1)), NA),
    c(paste(comp, '\n', outcome,'/', 'total n'),
      analysis$pre$rate2, paste(sum(analysis$pre$e2), " / ", sum(analysis$pre$n2)), NA),
    c("Odds Ratio (95% CI)", analysis$pre$or, 
      paste(formatC(analysis$exp$pred, format='f', digits =2), 
            " (",formatC(analysis$exp$ci.lb, format='f', digits=2),
            "-", formatC(analysis$exp$ci.ub, format='f', digits=2), ")"), NA),
    c("Weight(%)", paste(formatC(weights(analysis$ma), format='f', digits = 1),'%'), NA, NA)),
       b <- cbind( 
        c("Author", analysis$pre$author),
        c(paste(int, '\n', outcome,'/','total n'),analysis$pre$rate1),
        c(paste(comp, '\n', outcome,'/', 'total n'),analysis$pre$rate2),
        c("Odds Ratio (95% CI)", analysis$pre$or)))
  
  b <- as_tibble(b, .name_repair = "unique")
  b <- b %>% add_row(.before = 2)
  b <- b %>% add_row(.before = 2)
  b <- b %>% add_row(.after = nrow(b))  
  
  ifelse(nstudies>1, b <- b %>% add_row(.before = nrow(b)-2), NA)

  
  ifelse(nstudies > 1,
        (c <- structure(list(
          mean = c(rep(NA, 3), analysis$pre$est, NA, analysis$exp$pred,NA, NA),
          lower = c(rep(NA, 3), analysis$pre$ci.lb, NA, analysis$exp$ci.lb, NA, NA),
          upper = c(rep(NA, 3), analysis$pre$ci.ub, NA, analysis$exp$ci.ub, NA, NA)),
          .Names = c("mean", "lower", "upper"),
          row.names = c(NA, -1L*nrow(b)),
          class = "data.frame")),
        (c <- structure(list(
          mean = c(rep(NA, 3), analysis$pre$est, NA, NA),
          lower = c(rep(NA, 3), analysis$pre$ci.lb, NA, NA),
          upper = c(rep(NA, 3), analysis$pre$ci.ub, NA, NA)),
          .Names = c("mean", "lower", "upper"),
          row.names = c(NA, -1L*nrow(b)),
          class = "data.frame")))

  c <- as_tibble(c)
  
  list(b = b, c = c)
} 
  

plotor_single_studies <- function(words, numbers, xtick, sizes, 
                                  bolding, aligning, fpPosition){
  
  forestplot(words,
             graph.pos = fpPosition,
             zero = 1,
             numbers,
             new_page = TRUE,
             colgap = unit(5, "mm"),
             hrzl_lines = list("3" = gpar (lwd=1, columns=c(1:(ncol(words)+1)), col="black")),
             lineheight=unit(0.7,'cm'),
             boxsize = sizes,
             line.margin = 2,
             is.summary = bolding,
             align = aligning,
             ci.vertices = TRUE,
             txt_gp = fpTxtGp(label =gpar (cex=0.9), 
                              ticks = gpar(cex = 0.9, fontface="bold"),
                              summary = gpar(cex = 0.9),
                              xlab = gpar(cex=0.9)),
             xticks = xtick,
             xlog=TRUE,
             clip = c(0.2, 32),
             grid = xtick,
             lwd.xaxis = 1,
             lwd.ci = 2.2,
             lwd.zero = 2,
             graphwidth = unit(10,"cm"),
             col=fpColors(box="black",line="grey", zero = 'dodgerblue4', axes="grey20", summary="black"))
  
  
}  
```


<h2>Right robotic vs Left robotic Continuos analysis: Mean Difference (MD) 95%CI</h2>


<details>

<summary><b>Figure A -</b> Blood Loss (mL), Legnth of Hospital Stay (days) and Operative Time (min)</summary>

<br>
```{r overall_graph_REML, echo=FALSE, message=FALSE, warning=FALSE, fig.height = 9.9, fig.width = 15.5}
robotic_cont <- read.csv("data_extraction_continous.csv")
names(robotic_cont)[2] <- "author"
names(robotic_cont)[8] <- "n2"
names(robotic_cont)[9] <- "n1"
names(robotic_cont)[10] <- "mean_post2" #Change for control group
names(robotic_cont)[12] <- "sd_post2"   #Change for control group
names(robotic_cont)[13] <- "mean_post1" #Change for intervention group
names(robotic_cont)[15] <- "sd_post1"   #Change for intervention group


#Blood loss
bl <- subset(robotic_cont, outcome=='blood loss')
bl_md <- md(bl)
tbl_md <- table_md(analysis = bl_md, nstudies = 2, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Total blood loss (bl)', col = 'Mean (SD)', 
                    follow = FALSE)


#Operative time
ot <- subset(robotic_cont, outcome=='Operative time')
ot_md <- md(ot)
tot_md <- table_md(analysis = ot_md, nstudies = 2, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Operative time (min)', col = 'Mean (SD)', 
                    follow = FALSE)

#length of hospital stay
los <- subset(robotic_cont, outcome=='length of hospital stay')
los_md <- md(los)
tlos_md <- table_md(analysis = los_md, nstudies = 8, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Length of Hospital stay (min)', col = 'Mean (SD)', 
                    follow = FALSE)


title <- c('', paste0('Right robotic (N)'), "mean (SD)",
            paste0('Left robotic (N)'), "mean (SD)",
            paste0('Mean Difference (MD) 95%CI'), 
            paste0('Weights'))
  

sub_cont1 <- c('Blood Loss (mL)', rep(NA, 6))
sub_cont2 <- c('Legnth of Hospital Stay (days)', rep(NA, 6))
sub_cont3 <- c('Operative Time (min)', rep(NA, 6))


subtotal2 <- rbind(title, NA,
                   sub_cont1, tbl_md$b[-c(1:3, nrow(tbl_md$b)-3),], NA,
                   sub_cont2, tlos_md$b[-c(1:3, nrow(tlos_md$b)-3),], NA,
                   sub_cont3, tot_md$b[-c(1:3, nrow(tot_md$b)-3),])
subtotal2 <- as.matrix(subtotal2) #Sol: Error in assertAndRetrieveTidyValue(x, lower) : 
#You have not provided an argument and the data frame does not have a 'lower' column: V1, V2, V3, V4, V5, V6, V7

subtotal2_summary <- c(T, F, 
                           T, rep(F, nrow(tbl_md$b[-c(1:3, nrow(tbl_md$b)-3),])-3),T,T,F,F,
                           T, rep(F, nrow(tlos_md$b[-c(1:3, nrow(tlos_md$b)-3),])-3), T, T,F,F,
                           T, rep(F, nrow(tot_md$b[-c(1:3, nrow(tot_md$b)-3),])-3), T, T,F)

sizes_cont <- c(rep(NA,3), 0.006*(bl_md$pre$weights),1,
                rep(NA, 4), 0.03*(los_md$pre$weights),1,
                rep(NA, 4), 0.05*(ot_md$pre$weights),1,
                rep(NA, 2))

subtotal2n <- rbind(NA,NA,
                    NA, tbl_md$c[-c(1:3, nrow(tbl_md$c)-3),], NA,
                   NA, tlos_md$c[-c(1:3, nrow(tlos_md$c)-3),], NA,
                   NA, tot_md$c[-c(1:3, nrow(tot_md$c)-3),])


plot <- forestplot(subtotal2,
             graph.pos = 6,
             zero = 0,
             subtotal2n,
             new_page = TRUE,
             colgap = unit(5, "mm"),
             hrzl_lines = list("2" = gpar (lwd=1, columns=1:8, col="black")),
             lineheight=unit(0.7,'cm'),
             boxsize = sizes_cont,
             line.margin = 2,
             is.summary = subtotal2_summary,
             align = c('l', rep('c', 4), 'l', 'l'),
             ci.vertices = TRUE,
             txt_gp = fpTxtGp(label =gpar (cex=0.9), 
                              ticks = gpar(cex = 0.9, fontface="bold"),
                              summary = gpar(cex = 0.9),
                              xlab = gpar(cex=0.9)),
             xticks = c(-90,-45,0,30,50),
             xlog=FALSE,
             clip = c(-90,-45,30,50),
             grid = c(-90,-45,0,30,50),
             lwd.xaxis = 1,
             lwd.ci = 2.2,
             lwd.zero = 2,
             graphwidth = unit(8,"cm"),
             col=fpColors(box="black",line="grey", zero = 'dodgerblue4', axes="grey20", summary="black"))

plot

```

</details>

```{r dichotomous_analysis, echo=FALSE, message=FALSE, warning=FALSE}
robotic_dic <- read.csv("data_extraction_dicotomica.csv")
names(robotic_dic)[2] <- "author"
names(robotic_dic)[10] <- "e2" #Changed for Author's decisition
names(robotic_dic)[11] <- "e1" #Changed for Author's decisition
names(robotic_dic)[8] <- "n2" #Changed for Author's decisition
names(robotic_dic)[9] <- "n1" #Changed for Author's decisition

#Anastomotic leak
leak <- subset(robotic_dic, outcome=='Anastomotic leak')
leak <- leak[-1,]
leak_or <- or(leak)
tt_leak <- table_or(analysis = leak_or, nstudies = 3, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Anastomotic Leak')

#Conversion operation
conversion <- subset(robotic_dic, outcome=='Conversion operation')
conv_or <- or(conversion)
tt_conv <- table_or(analysis = conv_or, nstudies = 5, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Conversion Operation')

#Ileus
ileus <- subset(robotic_dic, outcome=='Ileus')
ileus_or <- or(ileus)
tt_ileus <- table_or(analysis = ileus_or, nstudies = 4, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Ileus')


#Readmission
readm <- subset(robotic_dic, outcome=='readmission')
readm <- readm[-1,]
readm_or <- or(readm)
tt_readm <- table_or(analysis = readm_or, nstudies = 4, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Readmission')

#Reoperation
resop <- subset(robotic_dic, outcome=='reoperation')
resop_or <- or(resop)
tt_resop <- table_or(analysis = resop_or, nstudies = 2, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Readmission')

#wound infection
w_inf <- subset(robotic_dic, outcome=='wound infection')
w_inf_or <- or(w_inf)
tt_w_inf <- table_or(analysis = w_inf_or, nstudies = 4, 
                    int = "Right robotic (n)", comp = "Left robotic (n)",
                    outcome = 'Wound Infection')

```



<h2>Right robotic vs Left robotic Dichotomous analysis: Odds Ratio (OR) 95%CI</h2>


<details>

<summary><b>Figure B -</b> Ileus, Conversion Operation, Anastomotic Leak, Reoperation. Wound infection, Readmission</summary>

<br>
```{r dic_graphs, echo=FALSE, fig.height=14, fig.width=14, message=FALSE, warning=FALSE}
title1 <- c('', paste0('Right robotic (n/N)'),
            paste0('Left robotic (n/N)'),
            paste0('Odds Ratio (OR) 95%CI'), 
            paste0('Weights'))
  

sub_dic1 <- c('Ileus', rep(NA, 4))
sub_dic2 <- c('Conversion Operation', rep(NA, 4))
sub_dic3 <- c('Anastomotic Leak', rep(NA, 4))
sub_dic4 <- c('Reoperation', rep(NA, 4))
sub_dic5 <- c('Wound infection', rep(NA, 4))
sub_dic6 <- c('Readmission', rep(NA, 4))


subtotal1 <- rbind(title1, NA,
                   sub_dic1, tt_ileus$b[-c(1:3, nrow(tt_ileus$b)-3),], NA,
                   sub_dic2, tt_conv$b[-c(1:3, nrow(tt_conv$b)-3),], NA,
                   sub_dic3, tt_leak$b[-c(1:3, nrow(tt_leak$b)-3),], NA,
                   sub_dic4, tt_resop$b[-c(1:3, nrow(tt_resop$b)-3),], NA,
                   sub_dic5, tt_w_inf$b[-c(1:3, nrow(tt_w_inf$b)-3),], NA,
                   sub_dic6, tt_readm$b[-c(1:3, nrow(tt_readm$b)-3),])
subtotal1 <- as.matrix(subtotal1)

sizes_dic <- c(rep(NA,3), 0.008*(ileus_or$pre$weights),1,
                rep(NA, 4), 0.02*(conv_or$pre$weights),1,
                rep(NA, 4), 0.008*(leak_or$pre$weights),1,
                rep(NA, 4), 0.008*(resop_or$pre$weights),1,
                rep(NA, 4), 0.008*(w_inf_or$pre$weights),1,
                rep(NA, 4), 0.008*(readm_or$pre$weights),1, NA, NA)
              
subtotal1_summary <- c(T, F, 
                           T, rep(F, nrow(tt_ileus$b[-c(1:3, nrow(tt_ileus$b)-3),])-3),T,T,F,F,
                           T, rep(F, nrow(tt_conv$b[-c(1:3, nrow(tt_conv$b)-3),])-3), T, T,F,F,
                           T, rep(F, nrow(tt_leak$b[-c(1:3, nrow(tt_leak$b)-3),])-3), T, T,F,F,
                           T, rep(F, nrow(tt_resop$b[-c(1:3, nrow(tt_resop$b)-3),])-3),T,T, F,F,
                           T, rep(F, nrow(tt_w_inf$b[-c(1:3, nrow(tt_w_inf$b)-3),])-3),T,T, F,F,
                           T, rep(F, nrow(tt_readm$b[-c(1:3, nrow(tt_readm$b)-3),])-3),T,T, F)
                           

subtotal1n <- rbind(NA,NA,
                    NA, tt_ileus$c[-c(1:3, nrow(tt_ileus$c)-3),], NA,
                   NA, tt_conv$c[-c(1:3, nrow(tt_conv$c)-3),], NA,
                   NA, tt_leak$c[-c(1:3, nrow(tt_leak$c)-3),], NA,
                   NA, tt_resop$c[-c(1:3, nrow(tt_resop$c)-3),], NA,
                   NA, tt_w_inf$c[-c(1:3, nrow(tt_w_inf$c)-3),], NA,
                   NA, tt_readm$c[-c(1:3, nrow(tt_readm$c)-3),])



plot1 <- forestplot(subtotal1,
             graph.pos = 4,
             zero = 1,
             subtotal1n,
             new_page = TRUE,
             colgap = unit(5, "mm"),
             hrzl_lines = list("2" = gpar (lwd=1, columns=1:6, col="black")),
             lineheight=unit(0.7,'cm'),
             boxsize = sizes_dic,
             line.margin = 2,
             is.summary = subtotal1_summary,
             align = c('l', rep('c', 2), 'l', 'l'),
             ci.vertices = TRUE,
             txt_gp = fpTxtGp(label =gpar (cex=0.9), 
                              ticks = gpar(cex = 0.9, fontface="bold"),
                              summary = gpar(cex = 0.9),
                              xlab = gpar(cex=0.9)),
             xticks = c(0.1,0.25, 0.50, 1, 2,4, 10),
             xlog=TRUE,
             clip = c(0.2, 10),
             grid = c(0.1,0.25, 0.50,  2,4,10),
             lwd.xaxis = 1,
             lwd.ci = 2.2,
             lwd.zero = 2,
             graphwidth = unit(10,"cm"),
             col=fpColors(box="black",line="grey", zero = 'dodgerblue4', axes="grey20", summary="black"))

plot1

```

</details>

